---
title: "RFM"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(ggplot2)
#library(stringr)
#library(DT)
library(tidyr)
library(knitr)
library(rmarkdown)
library("NbClust")
library("cluster")
library("factoextra")
library("magrittr")

```

```{r}
##Load Dataset
df_data <- fread('../data.csv')
glimpse(df_data)

```

```{r}
### Data Cleaning
#Delete all negative Quantity and Price. We also need to delete NA customer ID
df_data <- df_data %>% 
  mutate(Quantity = replace(Quantity, Quantity<=0, NA),
         UnitPrice = replace(UnitPrice, UnitPrice<=0, NA))

df_data <- df_data %>%
  drop_na()
```

```{r}
### Recode variables
#We should do some recoding and convert character variables to factors.

df_data <- df_data %>% 
  mutate(InvoiceNo=as.factor(InvoiceNo), StockCode=as.factor(StockCode), 
         InvoiceDate=as.Date(InvoiceDate, '%m/%d/%Y %H:%M'), CustomerID=as.factor(CustomerID), 
         Country=as.factor(Country))

df_data <- df_data %>% 
  mutate(total_dolar = Quantity*UnitPrice)

glimpse(df_data)
```


### Calculate RFM {.tabset}
To implement the RFM analysis, we need to further process the data set in by the following steps:

1. Find the most recent date for each ID and calculate the days to the now or some other date, to get the Recency data
2. Calculate the quantity of translations of a customer, to get the Frequency data
3. Sum the amount of money a customer spent and divide it by Frequency, to get the amount per transaction on average, that is the Monetary data.

```{r results='hold'}

df_RFM <- df_data %>% 
  group_by(CustomerID) %>% 
  summarise(recency=as.numeric(as.Date("2012-01-01")-max(InvoiceDate)),
            frequenci=n_distinct(InvoiceNo), monitery= sum(total_dolar)/n_distinct(InvoiceNo)) 

summary(df_RFM)

kable(head(df_RFM))
```
#### Recency
#Recency – How recently did the customer purchase?
```{r} 
hist(df_RFM$recency)
````



#### Frequency
#Frequency – How often do they purchase?
```{r} 
hist(df_RFM$frequenci, breaks = 50)
```



#### Monetary
#Monetary Value – How much do they spend?
```{r} 
hist(df_RFM$monitery, breaks = 50)
```
#Becouse the data is realy skewed, we use log scale to normalize
```{r}
df_RFM$monitery <- log(df_RFM$monitery)
hist(df_RFM$monitery)
```


#Plots
```{r}
#Plotting the data R Vs F
ggplot(data = df_RFM, mapping = aes(x=df_RFM$recency, y=df_RFM$monitery)) +
       geom_point(size=3)
```






### Clustering
```{r}
df_RFM2 <- df_RFM
row.names(df_RFM2) <- df_RFM2$CustomerID
df_RFM2$CustomerID <- NULL

df_RFM2 <- scale(df_RFM2)
summary(df_RFM2)
```

```{r}
d <- dist(df_RFM2)
c <- hclust(d, method = 'ward.D2')

plot(c)
```

#### Cut
```{r}
members <- cutree(c,k = 8)

members[1:5]
table(members)

aggregate(df_RFM[,2:4], by=list(members), mean)

```

```{r}
#Adi k mean implementation
#partitioning clustering
########################

pam.res <- pam(df_RFM, 3) #PAM clustering method
pam.res2 <- pam(df_RFM, 6) #PAM clustering method


fviz_cluster(pam.res) # Visualize
fviz_cluster(pam.res2) # Visualize

```

```{r}
#Ananya's k-means

aggregate(df_RFM[,2:4], by=list(members), mean)
```

```{r}
#new test
aggregate(df_RFM[,2:4], by=list(members), mean)

```

```{r}
#Demo code for heirachircal clustering
aggregate(df_RFM[,2:4], by=list(members), mean)
```

